---
title: "Working with `sf` geometry in this repo"
subtitle: "From per-slice ROIs to one outline per nucleus"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6, fig.height = 5)
```

## Why this note exists

`scripts/R/` turns Fiji's per-slice ROI outlines into polygons and groups them
into cellular features. Everything downstream — counting nuclei, drawing a QC
overview, measuring anything per feature — depends on understanding what kind of
object is being passed around. The types look interchangeable and are not.

The concrete task that prompted this: **one nucleus produces ~12 ROIs, one per
z-slice. The QC plot should draw one outline per nucleus, not twelve.**

> **Which R.** The spatial stack works under **R 4.6** on this machine, not
> under the rig default 4.4 — see `CLAUDE.local.md`. Pin the version without
> hardcoding a path:
>
> ```
> rig run -r 4.6 note/sf_geometry_primer.Rmd
> ```
>
> Knitting also needs `ggplot2`, `knitr` and `rmarkdown`, which are currently
> installed only under 4.4. Until they are in 4.6 too, the non-plotting chunks
> run but the document will not knit.

```{r libs, message=FALSE}
library(dplyr); library(tidyr); library(tibble); library(stringr)
library(sp); library(sf)
library(ggplot2)

repo <- rprojroot::find_root(rprojroot::has_file("VERSION"))
for (f in list.files(file.path(repo, "scripts/R"), full.names = TRUE)) source(f)
```

---

## 1. The three layers of `sf`

`sf` ("simple features") stacks three types. Nearly every confusion comes from
mistaking one for another.

| Layer | Class | What it is | Analogy |
|---|---|---|---|
| `sfg` | `"XY" "POLYGON" "sfg"` | **one** geometry | a single value |
| `sfc` | `"sfc_POLYGON" "sfc"` | a **column** of geometries, with a CRS | a vector |
| `sf`  | `"sf" "data.frame"` | a data frame **whose geometry column is registered** | a data frame |

```{r three-layers}
# sfg -- a single polygon. A polygon is a LIST of rings: the first is the
# outer boundary, any others are holes. Each ring must close (last == first).
square <- st_polygon(list(cbind(c(0, 4, 4, 0, 0),
                                c(0, 0, 4, 4, 0))))
class(square)

# sfc -- a column of them
polys <- st_sfc(square, square + c(6, 0))
class(polys)

# sf -- a data frame that KNOWS which column is the geometry
tbl <- st_sf(id = c("a", "b"), geometry = polys)
class(tbl)
```

Note `square + c(6, 0)`: `sfg` objects support arithmetic, which shifts them.

### The distinction that matters here

A tibble can *hold* an `sfc` column without *being* an `sf` object. It looks
identical when printed, and the geometry-aware methods do not fire.

```{r sf-vs-tibble}
plain <- tibble(id = c("a", "b"), geometry = polys)  # a tibble with a geometry column
real  <- st_as_sf(plain)                            # now registered

inherits(plain, "sf")
inherits(real,  "sf")
class(plain$geometry)   # identical in both cases -- this is why it is easy to miss
```

**This repo's functions return the plain kind.** `polygonize_roi_df()` and
`define_feature_group()` both build a tibble carrying an `sfc` column; neither
calls `st_as_sf()`. So a plain `tibble` in, a plain `tibble` out — and you must
convert before using any `sf` method.

---

## 2. Load the fixture and build the features

```{r load}
outline_path <- file.path(repo, "fixture/if_data/data/GRV_Position010_nucleus_outline.txt")
roi_df <- read.table(outline_path, header = TRUE, stringsAsFactors = FALSE) %>%
  as_tibble() %>%
  dplyr::select(-name)

glimpse(roi_df)
```

One row per polygon **vertex**: `roi`, `z`, `x`, `y`. `polygonize_roi_df()`
collapses the vertices of each ROI into one polygon.

```{r polygonize}
roi_pg_df <- polygonize_roi_df(roi_df)

nrow(roi_df)        # vertices
nrow(roi_pg_df)     # ROIs
class(roi_pg_df)    # <- a tibble, NOT sf
```

`define_feature_group()` then decides which ROIs across z belong to the same
object, by testing x-y intersection between ROIs within `max_z_dist` slices of
each other and taking connected components of that graph.

```{r features}
feature_df <- define_feature_group(
  roi_df,
  roi_regex   = "^nucleus",
  feature_prefix         = "nucleus_",
  invalid_feature_prefix = "invalid_",
  max_z_dist = 3,     # ROIs up to 3 slices apart may be linked
  min_z_span = 5      # a feature must appear on >= 5 distinct slices
)

table(str_extract(feature_df$feature_id, "^[a-z]+"))
```

65 ROIs were assigned to real nuclei; 5 fell into `invalid_` groups that did not
span enough slices. Still one row per ROI — the grouping is recorded in
`feature_id`, the geometry is untouched.

---

## 3. The union: one outline per nucleus

This is the answer to the original question, and it is one line.

```{r union}
nucleus_sf <- feature_df %>%
  dplyr::filter(str_detect(feature_id, "^nucleus_")) %>%
  st_as_sf()                                    # <-- REQUIRED, see section 1

nucleus_union <- nucleus_sf %>%
  group_by(feature_id) %>%
  summarise(n_roi = n(),
            z_min = min(z),
            z_max = max(z),
            .groups = "drop")

nucleus_union
```

**`summarise()` on an `sf` object unions the geometries of each group**, because
`sf`'s method defaults to `do_union = TRUE`. You get the merged outline for free
alongside whatever per-group statistics you list. 65 ROIs became 6 nuclei.

```{r union-checks}
table(as.character(st_geometry_type(nucleus_union)))
all(st_is_valid(nucleus_union))
round(as.numeric(st_area(nucleus_union)), 1)     # square microns
```

All six collapse to a single `POLYGON` — the per-slice ROIs of one nucleus
overlap enough that their union has no gap.

### What the union actually buys us

Fiji's `Overview.groovy` `merged` mode unions ROIs **in the flattened 2D
projection**, so anything overlapping in x-y becomes one outline regardless of
z. R groups by `feature_id`, which was built from z-aware overlap. They disagree,
and it is easy to see why:

```{r overlap-check}
ov <- st_intersects(nucleus_union, nucleus_union, sparse = FALSE)
diag(ov) <- FALSE
sum(ov) / 2                                  # overlapping pairs in x-y

w <- which(ov, arr.ind = TRUE)
unique(t(apply(w, 1, function(i) sort(nucleus_union$feature_id[i]))))
```

`nucleus_1` (slices 1–5) and `nucleus_6` (slices 24–35) overlap in x-y but are
30 slices apart. R keeps them separate; Fiji's projection merges them. **That is
the whole of the 5-vs-6 discrepancy** — Fiji's overlay is a picture, not a count.

---

## 4. `do_union = FALSE` is a different thing entirely

Reach for it by accident and the result still plots, still has six rows, and is
wrong.

```{r no-union}
not_unioned <- nucleus_sf %>%
  group_by(feature_id) %>%
  summarise(n_roi = n(), .groups = "drop", do_union = FALSE)

table(as.character(st_geometry_type(not_unioned)))
```

`MULTIPOLYGON`: all twelve per-slice rings kept side by side in one row, merely
*collected* rather than merged. Drawn, it is twelve nested outlines, not one.

```{r vertex-counts}
data.frame(
  feature  = nucleus_union$feature_id,
  unioned  = sapply(seq_len(nrow(nucleus_union)), function(i) nrow(st_coordinates(nucleus_union$geometry[i]))),
  collected = sapply(seq_len(nrow(not_unioned)),  function(i) nrow(st_coordinates(not_unioned$geometry[i])))
)
```

---

## 5. Why `plot_outline_topView()` cannot draw this

`plot_outline_topView()` goes through `pg_2_coord_df()`, which flattens geometry
to plain `x`/`y` columns and hands them to `geom_polygon(aes(group = roi))`.
That works for the per-slice ROIs it was written for — each is one simple ring.

It breaks on a union, for a structural reason:

```{r coords}
xy <- st_coordinates(nucleus_union$geometry[1])
head(xy, 3)
colnames(xy)
```

`st_coordinates()` returns **`L1` and `L2` alongside X and Y**: `L1` numbers the
ring within a polygon (1 = exterior, 2+ = holes) and `L2` numbers the polygon
within a MULTIPOLYGON. `pg_2_coord_df()` keeps only X and Y, so:

* a **hole** becomes another loop of the same outline, drawn as if it were part
  of the boundary — the ring and its interior join into one tangled shape;
* a **MULTIPOLYGON** has all its parts concatenated into a single `group`, so
  `geom_polygon` connects the last vertex of one part to the first of the next.

Neither errors. Both draw something plausible.

### Use `geom_sf`

`geom_sf` consumes the geometry column directly and understands rings and
multi-part geometries, so nothing has to be flattened.

```{r plot-union, fig.height=5}
plot_features_topView(nucleus_sf, nucleus_union, color_by = "feature_id") +
  labs(title = "One outline per nucleus (z-aware union)")
```

`plot_features_topView()` is the library function in
`scripts/R/plot_outline_topView.r`. It also handles the y-axis problem
described in section 6 — see there for why that is not a one-liner.

Side by side with every per-slice ROI, which is what Fiji's `all` mode shows:

```{r plot-compare, fig.width=10, fig.height=5}
per_slice <- nucleus_sf %>% mutate(panel = "per-slice ROIs (65)")
unioned   <- nucleus_union %>% mutate(panel = "unioned features (6)")
bb <- st_bbox(nucleus_sf); y_ref <- unname(bb["ymin"] + bb["ymax"])

bind_rows(
  dplyr::select(per_slice, feature_id, panel, geometry),
  dplyr::select(unioned,   feature_id, panel, geometry)
) %>%
  flip_y_image(y_ref = y_ref) %>%
  ggplot() +
  geom_sf(aes(colour = feature_id), fill = NA, linewidth = 0.4) +
  facet_wrap(~ panel) +
  coord_sf() +
  scale_y_continuous(breaks = flip_y_breaks(y_ref), labels = flip_y_labels(y_ref)) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

---

## 6. Gotchas worth remembering

**`summarise()` silently does nothing geometric on a non-`sf` tibble.** No
error; the geometry column is simply dropped from the result. Always
`st_as_sf()` first.

**`st_as_sf()` needs to know the column.** It finds a lone `sfc` column
automatically; with more than one, name it: `st_as_sf(df, sf_column_name = "geometry")`.

**`dplyr::select()` on an `sf` object keeps the geometry whether you ask or
not** — it is "sticky". To actually drop it use `st_drop_geometry()`.

**There is no CRS here, and that is correct.** These are microns in image space,
not positions on Earth. `st_area()` therefore returns plain numbers in µm², and
`sf` may warn about assuming planar coordinates — which is exactly what we want.
Never set a geographic CRS on this data.

**`st_area()` returns a `units` object** when a CRS is set, and a bare numeric
otherwise. Wrap in `as.numeric()` before comparing to a plain threshold if you
want the same code to work either way.

**Validity before union.** `st_union()` on a self-intersecting ring can produce
garbage rather than an error. Fixture outlines are all valid, but traced ROIs
from noisier data may not be:

```{r validity}
all(st_is_valid(nucleus_sf))
# if this is ever FALSE:  nucleus_sf <- st_make_valid(nucleus_sf)
```

**⚠️ `coord_sf()` silently ignores `scale_y_reverse()`.** This is the nastiest
one here. Image coordinates put the origin at the top-left with y increasing
downward; plots do the opposite, so every image plot needs a flip. The obvious
spelling does not work:

```{r yflip-demo}
sq <- st_sf(id = "L", geometry = st_sfc(st_polygon(list(
  cbind(c(0, 10, 10, 6, 6, 0, 0), c(0, 0, 2, 2, 8, 8, 0))))))

pp <- function(p) { r <- ggplot_build(p)$layout$panel_params[[1]]; r$y_range }
pp(ggplot(sq) + geom_sf() + coord_sf())                      # plain
pp(ggplot(sq) + geom_sf() + scale_y_reverse() + coord_sf())  # identical
pp(ggplot(sq) + geom_sf() + coord_sf(ylim = c(8, 0)))        # identical
```

All three render the same picture. No warning, no error — the QC panel is just a
vertical mirror of the Fiji PNG, which looks entirely reasonable until you try to
compare them. (`coord_fixed()` is not an escape either: `geom_sf` requires
`coord_sf`.)

The only reliable fix is to **flip the geometry** and relabel the axis, which is
what `flip_y_image()` / `flip_y_breaks()` / `flip_y_labels()` in
`scripts/R/plot_outline_topView.r` do:

```{r yflip-fix, eval=FALSE}
y_ref <- unname(st_bbox(x)["ymin"] + st_bbox(x)["ymax"])
ggplot(flip_y_image(x, y_ref)) + geom_sf() + coord_sf() +
  scale_y_continuous(breaks = flip_y_breaks(y_ref), labels = flip_y_labels(y_ref))
```

Pass an explicit `y_ref` (the image height) whenever several panels must share
one extent — otherwise each panel flips about its own bounding box and they
stop being comparable.

---

## 7. Cheat sheet

```{r cheatsheet, eval=FALSE}
st_as_sf(df)                  # tibble-with-geometry  ->  real sf object
st_sf(id = x, geometry = g)   # build one from scratch
st_drop_geometry(sf)          # sf -> plain tibble (select() will not do it)

st_geometry_type(sf)          # POLYGON / MULTIPOLYGON / ...
st_is_valid(sf)               # self-intersections?
st_make_valid(sf)             # repair them

st_area(sf)                   # per row
st_centroid(sf)               # per row
st_bbox(sf)                   # xmin/ymin/xmax/ymax of the whole layer

st_intersects(a, b, sparse = FALSE)   # logical matrix: do they touch?
st_intersection(a, b)                 # the overlapping geometry itself
st_union(sfc)                         # merge a whole column into one geometry

group_by(sf, g) %>% summarise(...)    # per-group union + statistics  <- the workhorse
                                      # do_union = FALSE collects instead of merging

geom_sf(aes(colour = f), fill = NA)   # draws rings and multi-parts correctly
st_coordinates(g)                     # X, Y, L1 (ring), L2 (part)

# this repo (scripts/R/plot_outline_topView.r)
union_features(sf)                    # per-slice ROIs -> one outline per feature
plot_features_topView(roi_sf, union)  # the QC plot, y already in image orientation
flip_y_image(sf, y_ref)               # coord_sf ignores scale_y_reverse(); flip the data
```
