---
title: "Checking Nr5a2 signals"
author: "Author: Siwat (Chad) Ruangroengkulrith (Tachibana Lab)"
date: "Created: 06/04/2024"
output: 
  html_notebook:
    theme: flatly
    highlight: pygments
    toc: true
    # toc_depth: 2
    # toc_float: true
    # code_folding: hide
---
```{r, echo=FALSE, results='asis'}
# cat(format(Sys.time(), "%a %b %d %X %Y"))
cat("Last Updated and Executed:", format(Sys.time(), "%d %B %Y %X"))
```

```{r}
addDateSuffix <- TRUE
if(addDateSuffix){
  date_suffix <- format(Sys.time(), "_%Y%m%d")
}else{
  date_suffix <- ""
}
```

```{r}
suppressPackageStartupMessages({
  # Generic
  library(dplyr)
  library(tidyr)
  library(tibble)
  library(stringr)
  library(readxl)
  
  # Spatial analysis
  library(sp)
  library(sf)
  
  # plot function
  library(ggplot2)
  library(ggbeeswarm)
  library(plotly)
  library(ggplotify)
  library(cowplot)
  library(ggpubr)
})
```


```{r}
wd <- "/Users/chad/Lab/Projects/Maternal_Nr5a2/IF/20240327_timecourse_nr5a2"
script_dir <- "/Users/chad/Lab/dev/embryo_IF_analyzer/scripts/R"

RData_dir <- paste0(wd, "/data")
if(!dir.exists(RData_dir)){dir.create(RData_dir)}
```

```{r}
source(paste0(script_dir, "/brewer_pal_2.r"))
source(paste0(script_dir, "/read_fiji_result.r"))
source(paste0(script_dir, "/FnGroup_roi_2_polygons.r"))
source(paste0(script_dir, "/find_ROI_z_intersect.r"))
source(paste0(script_dir, "/define_feature_group.r"))
source(paste0(script_dir, "/find_overlap_roi_features.r"))

# Plot functions
source(paste0(script_dir, "/plot_outline_topView.r"))
```


# Read test data

***************************************************************************************************

```{r}
# Original
data_dir <- "/Volumes/pool-toti-imaging/Alina/LeicaStellaris/27032024_Nr5a2_timecourse_1hRT/SR/raw_measurements"
# local copy
data_dir <- "/Users/chad/Lab/Projects/Maternal_Nr5a2/IF/20240327_timecourse_nr5a2/raw_measurements"
plot_dir=paste0(data_dir, "/plot")
if(!dir.exists(plot_dir)){dir.create(plot_dir)}

channel_map <- c("DAPI", "H3K27ac", "Nr5a2")
channel_order <- c("Nr5a2", "H3K27ac", "DAPI")

pos_info <- read_xlsx(paste0(data_dir, "/../pos_info.xlsx"), sheet="sample") %>% 
  mutate(condition = str_replace(condition, "\\\\n", "\n"))
# pos_param_info <- read_xlsx(paste0(data_dir, "/../pos_info.xlsx"), sheet="parameter") 

background_raw_df <- read_xlsx(paste0(data_dir, "/../pos_info.xlsx"), sheet = "background")
```

```{r}
# pos_info$genotype %>% unique
genotype_color_map <- c(`fl/fl`="darkorchid4", KO = "firebrick")

condition_order <- c("fl/fl", "KO", "fl/fl\nNo_primary", "KO\nNo_primary")
```

# Calculating background signal

***************************************************************************************************

> The recommended way fro normalizing the signal is: `(Mean[ROI] x Area[ROI]) - (Mean[BG] x Area[BG])`. In other words: `(Mean[ROI] - Mean[BG]) x Area[ROI]`

```{r}
background_df <- background_raw_df %>% 
  dplyr::rename_all(.funs=function(x){tolower(str_remove_all(x, "\\.+$"))}) %>% 
  dplyr::select(-position) %>% 
  mutate(pos = str_extract(label, "[Pp]osition\\d+"),
         z = as.numeric(str_extract(label, "(?<=z\\:)\\d+"))) %>% 
  group_by(pos, ch, type) %>% 
  reframe(n_z = length(unique(z)),
          sum_intDen = sum(intden),
          sum_area = sum(area),
          mean_area = mean(area),
          signal = sum_intDen/sum_area,
          mean_signal = mean(mean)) %>% 
  mutate(channel = channel_map[as.integer(ch)]) %>% 
  left_join(., dplyr::select(pos_info, pos=position, genotype, celltype, hpf, condition), by="pos")

background_area_df <- background_df %>% 
  dplyr::select(pos, type, mean_area) %>% 
  unique() %>% 
  tidyr::spread(key="type", value="mean_area") %>% 
  dplyr::rename(bg_mean_area = background, cyto_mean_area=cytoplasm)
```

## Compare raw background signal
```{r}
background_df %>% 
  ggplot(aes(x = condition, y=signal)) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.5) +
  facet_grid(channel~type, scales="free_y") +
  labs(title = "Raw signal in two types of background",
       x="", y="Signal (Total IntDen / Total Area) (A.U.)") +
  theme_bw()
```


Normalizing the signal

```{r}
background_df_2 <- background_df %>%
  dplyr::select(pos, channel, type, signal) %>% 
  spread(key = "type", value="signal") %>% 
  dplyr::rename(bg_mean_signal = background, cyto_mean_signal=cytoplasm) %>% 
  # Add area column:
  left_join(., background_area_df, by="pos")
```


```{r}
background_norm_df <- background_df_2 %>% 
  mutate(norm_cyto_signal_1 = cyto_mean_signal/bg_mean_signal,
         norm_cyto_signal_2 = cyto_mean_signal - bg_mean_signal,
         norm_cyto_signal_3 = (cyto_mean_signal - bg_mean_signal)*cyto_mean_area) %>% 
  left_join(., dplyr::select(pos_info, pos=position, genotype, celltype, hpf, condition), by="pos")
```

```{r}
background_norm_df %>% 
  ggplot(aes(x=condition, y=log2(norm_cyto_signal_1))) +
  # ggplot(aes(x=condition, y=norm_cyto_signal_2)) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.5) +
  facet_grid(channel~., scales="free_y") +
  labs(title = "Normalized cytoplasmic signal over the background",
       x="", y="Signal log2(cytoplasm / background) (A.U.)") +
  theme_bw()

background_norm_df %>% 
  ggplot(aes(x=condition, y=log2(norm_cyto_signal_2))) +
  # ggplot(aes(x=condition, y=norm_cyto_signal_2)) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.5) +
  facet_grid(channel~., scales="free_y") +
  labs(title = "Normalized cytoplasmic signal over the background",
       x="", y="Signal (cytoplasm - background) (A.U.)") +
  theme_bw()

background_norm_df %>% 
  ggplot(aes(x=condition, y=log2(norm_cyto_signal_3))) +
  # ggplot(aes(x=condition, y=norm_cyto_signal_2)) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.5) +
  facet_grid(channel~., scales="free_y") +
  labs(title = "Normalized cytoplasmic signal over the background",
       x="", y="Signal (cytoplasm - background) * Area (A.U.)") +
  theme_bw()
```


# Read nucleus IF signal 

***************************************************************************************************

# Read data

```{r}
# nuc_data_path <- paste0(RData_dir, "/nuc_raw_p1_113.RData")
nuc_data_path <- paste0(RData_dir, "/nuc_raw_p114_141.RData")

pos_range <- 114:142
# pos_range <- 82:142
pos_id_vec <- paste0("Position", str_pad(pos_range, width=3, side="left", pad="0"))
prefered_suffix <- "byCH1_Huang2"
prefered_suffix_regex <- c("(byCH1)")
```

```{r}
do_read_nuc_data <- TRUE
if(!is.null(nuc_data_path)){
  if(file.exists(nuc_data_path)){
    load(nuc_data_path)
    do_read_nuc_data <- FALSE
  }
}

# Read data =======================================================================================
if(do_read_nuc_data){
  all_feature_df <- tibble()
  all_res_df <- tibble()
  all_nucleo_nuc_ovl_df <- tibble()
  
  for(i in seq_along(pos_id_vec)){
    # Reading file ----------------------------------------------------------------------------------
    # pos_id <- "Position115"
    pos_id <- pos_id_vec[i]
    
    # NB: each FEATURE now has its own pair of files. The old Fiji macro wrote
    #     nucleus AND nucleolus ROIs into the nucleolus files, so reading
    #     "*_nucleolus_*" alone was enough and the nuclei came along for free.
    #     The Groovy scripts scope each feature to its own files, so that same
    #     read returns nucleolus ROIs only -- nuc_outline_df below comes back
    #     empty, every nucleus is lost, and the analysis yields nothing while
    #     only warning. Read both features and combine them.
    resolve_feature_path <- function(feature, kind){
      hits <- list.files(data_dir,
                         pattern=paste0(pos_id, ".+", feature, "_", kind, "\\.txt$"))
      if(length(hits) > 1){
        # Use the prefered suffix
        hits <- subset(hits, str_detect(hits, prefered_suffix_regex))[1]
      }
      if(length(hits) != 1 || is.na(hits)) return(NA_character_)
      paste0(data_dir, "/", hits)
    }
    
    feature_paths <- list(
      nucleus   = c(outline = resolve_feature_path("nucleus", "outline"),
                    res     = resolve_feature_path("nucleus", "res")),
      nucleolus = c(outline = resolve_feature_path("nucleolus", "outline"),
                    res     = resolve_feature_path("nucleolus", "res"))
    )
    
    # Old-format data keeps both features in the nucleolus files. Detect that
    # rather than assume either layout: if there is no separate nucleus file,
    # fall back to reading the nucleolus pair alone, which still carries both.
    split_files <- !is.na(feature_paths$nucleus["outline"]) &&
                   !is.na(feature_paths$nucleus["res"])
    read_features <- if(split_files) c("nucleus", "nucleolus") else "nucleolus"
    
    if(any(is.na(unlist(feature_paths[read_features])))){
      # Skip to the next iteration of the loop
      cat(pos_id, " - skip\n")
      next
    }
    cat(pos_id, if(split_files) "(split) " else "(combined) ")
    
    outline_df <- tibble(); res_df <- tibble()
    for(feat in read_features){
      outline_df <- as_tibble(read.table(feature_paths[[feat]]["outline"],
                                         header=TRUE, stringsAsFactors=FALSE)) %>%
        bind_rows(outline_df, .)
      res_df <- suppressWarnings(read_fiji_result(feature_paths[[feat]]["res"])) %>%
        bind_rows(res_df, .)
    }
    outline_df$type <- str_extract(outline_df$roi, "^[:alpha:]+")
    res_df <- mutate(res_df, pos=pos_id)
    res_df$type <- str_extract(res_df$roi, "^[:alpha:]+")
    
    # Fail loudly on the assumption that actually matters. Without this the
    # nucleus-free case below only warning()s -- its `next` is commented out --
    # and the run completes having silently produced nothing.
    if(!any(outline_df$type == "nucleus")){
      stop("No nucleus ROIs for ", pos_id,
           ": read ", paste(read_features, collapse=" + "),
           ". Expected a *_nucleus_outline.txt beside the nucleolus files.")
    }
    
    # Filter valid ROI ------------------------------------------------------------------------------
    min_circ <- pos_info %>% 
      dplyr::filter(position == pos_id) %>% 
      dplyr::pull(min_circularity)
    if(length(min_circ)!=1){
      min_circ <- 0.7 # Default value
    }
    
    valid_roi <- res_df %>% 
      dplyr::filter(circ>min_circ) %>%
      dplyr::pull(roi) %>% unique()
    
    # outline_df <- dplyr::filter(outline_df, roi %in% valid_roi)
    outline_df <- outline_df %>% 
      mutate(include = roi %in% valid_roi)
    
    nuc_outline_df <- dplyr::filter(outline_df, type == "nucleus")
    nucleo_outline_df <- dplyr::filter(outline_df, type == "nucleolus")
    
    if(sum(nuc_outline_df$include)==0){
      warning("No ROI pass circularity filter\n")
      # next
    }
    
    # Detect nucleus/nucleolus ----------------------------------------------------------------------
    nuc_df <- nuc_outline_df %>% 
      # dplyr::filter(type == "nucleus") %>% 
      define_feature_group(., pre_roi_filter_colname="include", 
                           roi_regex="^nucleus", feature_prefix="nucleus_", 
                           roi_area_range = c(200, 600),
                           min_intersect_ratio=0.90,
                           # max_z_dist=2, min_z_span=3, min_avg_area=100, # Setting for 2C onward
                           max_z_dist=30, min_z_span=3, min_avg_area=200, # Setting for zygote
                           invalid_feature_prefix="invalid_") %>% 
      mutate(type = "nucleus", pos = pos_id)
    
    n_nuc <- length(unique(dplyr::pull(dplyr::filter(nuc_df, str_detect(feature_id, "^nuc")), feature_id)))
    if(n_nuc>0){
      nuc_df <- mutate(nuc_df, nucleus_id = str_extract(feature_id, "^nucleus_\\d+"))
    }else{
      nuc_df$nucleus_id <- NA
    }
    
    if(n_nuc==0){
      warning("Couldn't detect nucleus, skip\n")
      # next
      # Maybe we should done something else
    }
    
    if(sum(nucleo_outline_df$include)>0){
      nucleo_df <- nucleo_outline_df %>% 
        # dplyr::filter(type == "nucleolus") %>% 
        define_feature_group(., pre_roi_filter_colname="include", 
                             roi_regex="^nucleolus", feature_prefix="nucleolus_", 
                             min_intersect_ratio=0.75,
                             invalid_feature_prefix="invalid_", 
                             max_z_dist=1, min_z_span=2) %>% 
        mutate(type = "nucleolus", pos = pos_id)
      n_nucleo <- length(unique(dplyr::pull(dplyr::filter(nucleo_df, str_detect(feature_id, "^nuc")), feature_id)))
    }else{
      nucleo_df <- tibble()
      n_nucleo=0
    }
    
    # split the results based on the compartment types
    # Which nucleolus belong to which nucleus
    if((n_nuc>0) & (n_nucleo>0)){
      nucleo_nuc_ovl_df <- find_overlap_roi_features(feature_df_1=nucleo_df, 
                                                     feature_df_2=nuc_df, 
                                                     feature_1_regex="nucleolus", feature_2_regex="nucleus", 
                                                     # somehow doesn't work when using min_intersect_ratio=1.0
                                                     min_intersect_ratio=0.9, min_ratio_roi_overlap=1) %>% 
        mutate(pos = pos_id)
      if(nrow(nucleo_nuc_ovl_df) > 0){
        feature_ovl_map <- dplyr::pull(nucleo_nuc_ovl_df, feature_id_2, name=feature_id_1)
        # Assigning nucleus_id field
        
        nucleo_df <- mutate(nucleo_df, nucleus_id = feature_ovl_map[feature_id])
      }else{
        nucleo_df$nucleus_id <- NA
      }
    }else{
      nucleo_df$nucleus_id <- NA
    }
    
    # Combine
    feat_df <- unique(rbind(nuc_df, nucleo_df))
    n_dup_roi <- feat_df$roi %>% duplicated() %>% sum()
    if(n_dup_roi>0){paste0("Duplicated ROI found: \n", pos_id, "; (i=", i, ")")}
    
    # Add feature_id to res_df
    res_df <- left_join(res_df, dplyr::select(feat_df, roi, feature_id, nucleus_id), by="roi")
    
    # Combine to the data ---------------------------------------------------------------------------
    if(nrow(all_res_df)==0){
      all_feature_df <- feat_df
      all_res_df <- res_df
      all_nucleo_nuc_ovl_df <- nucleo_nuc_ovl_df
    }else{
      # NB: Consider improve this part, this only work well when things went smoothly with the first position.
      all_feature_df <- add_row(all_feature_df, feat_df)
      all_res_df <- add_row(all_res_df, dplyr::select(res_df, any_of(colnames(all_res_df))))
      all_nucleo_nuc_ovl_df <- add_row(all_nucleo_nuc_ovl_df, nucleo_nuc_ovl_df)
    }
  }
  
  save(list=c("all_feature_df", "all_res_df", "all_nucleo_nuc_ovl_df"), file=nuc_data_path)
  cat("\nDone\n")
}
```


## Check pre-determined excluded nucleus

```{r}
if("exclude_nuc" %in% colnames(pos_info)){
  excl_nuc_df <- pos_info %>% 
    dplyr::filter(!is.na(exclude_nuc)) %>% 
    dplyr::select(pos=position, exclude_nuc)
  
  excl_nuc_list <- str_split(str_remove_all(excl_nuc_df$exclude_nuc, "\\s+"), ";") %>% 
    setNames(nm=excl_nuc_df$pos)
  
  # Get unique vector of excluded Nuc
  excl_nuc_vec <- c()
  for(i in seq_along(excl_nuc_list)){
    cur_id = paste0(names(excl_nuc_list)[i], "::", excl_nuc_list[[i]])
    excl_nuc_vec <- c(excl_nuc_vec, cur_id)
  }
  
  # Turn excluded nuc id into "invalid"
  all_unq_nuc_id <- all_feature_df %>% 
    transmute(unq_nuc_id = paste0(pos, "::", feature_id)) %>% 
    dplyr::pull(unq_nuc_id)
  invalid_flag <- all_unq_nuc_id %in% excl_nuc_vec
  
  all_feature_df$feature_id[invalid_flag] <- "invalid"
  
  # Change res_df
  all_res_df
  all_unq_nuc_id <- all_res_df %>% 
    transmute(unq_nuc_id = paste0(pos, "::", feature_id)) %>% 
    dplyr::pull(unq_nuc_id)
  invalid_flag <- all_unq_nuc_id %in% excl_nuc_vec
  
  all_res_df$feature_id[invalid_flag] <- "invalid"
  
}
```


# Ploting feature outlines

***************************************************************************************************

```{r, eval=FALSE, echo=FALSE}
unq_pos_vec <- unique(all_feature_df$pos) 
extend_factor <- 0.1
line_alpha=0.9
line_width=0.1

exclude_col_map <- c(invalid = "grey40", failed_ROI="gray50", other_feature = "gray75")

p_top_list <- list()
for(i in seq_along(unq_pos_vec)){
  pos_id <- unq_pos_vec[i]
  cat(unq_pos_vec[i], ",")
  feat_df <- all_feature_df %>% 
    dplyr::filter(pos == unq_pos_vec[i]) %>% 
    mutate(feature_id = str_replace(str_replace(feature_id, "^invalid.+", "invalid"), "^fail.+", "failed_ROI"))
  
  feat_coord_df <- pg_2_coord_df(feat_df)
  lim_list <- list(xmin=min(feat_coord_df$x, na.rm=TRUE)*(1-extend_factor), 
                   xmax=max(feat_coord_df$x, na.rm=TRUE)*(1+extend_factor),
                   ymin=min(feat_coord_df$y, na.rm=TRUE)*(1-extend_factor), 
                   ymax=max(feat_coord_df$y, na.rm=TRUE)*(1+extend_factor))
    
  # plot nucleus 
  unq_nuc_id <- feat_df %>% 
    dplyr::filter(str_detect(feature_id, "nucleus")) %>% 
    dplyr::pull(feature_id) %>% unique()
  unq_nucleo_id <- feat_df %>% 
    dplyr::filter(str_detect(feature_id, "nucleolus")) %>% 
    dplyr::pull(feature_id) %>% unique()
  
  n_nuc <- length(unq_nuc_id)
  n_nucleo <- length(unq_nucleo_id)
  
  nuc_col_map <- brewer_pal_2(n_nuc, pal_name="Paired")
  if(!is.null(nuc_col_map)){names(nuc_col_map) <- unq_nuc_id}
  nuc_col_map <- c(nuc_col_map, exclude_col_map)
  
  nucleo_col_map <- brewer_pal_2(n_nucleo, pal_name="Paired")
  if(!is.null(nucleo_col_map)){names(nucleo_col_map) <- unq_nucleo_id}
  nucleo_col_map <- c(nucleo_col_map, exclude_col_map)
  
  suppressWarnings({
    p1 <- feat_df %>% 
      dplyr::mutate(feature_id = if_else(type!="nucleus", "other_feature", feature_id)) %>%
      # dplyr::pull(feature_id) %>% table()
      plot_outline_topView(color_by="feature_id", 
                           color_map=nuc_col_map,
                           line_alpha=line_alpha, line_width=line_width) +
      coord_fixed(ratio=1, xlim=c(lim_list$xmin, lim_list$xmax), ylim=c(lim_list$ymin, lim_list$ymax)) +
      scale_y_reverse() +
      labs(title=paste0(pos_id, " - Nucleus (all)"))
    
    plot_df <- feat_df %>% 
      dplyr::filter(type=="nucleus", str_detect(feature_id, "^nuc"))
    if(nrow(plot_df) > 0){
      p2 <- plot_df %>% 
      plot_outline_topView(color_by="feature_id", 
                           color_map=nuc_col_map,
                           line_alpha=line_alpha, line_width=line_width) +
      coord_fixed(ratio=1, xlim=c(lim_list$xmin, lim_list$xmax), ylim=c(lim_list$ymin, lim_list$ymax)) +
        scale_y_reverse() +
      labs(title=paste0(pos_id, " - Nucleus (feature only)"))
    }else{
      p2 <- ggplot() +
        coord_fixed(ratio=1, xlim=c(lim_list$xmin, lim_list$xmax), ylim=c(lim_list$ymin, lim_list$ymax)) +
        scale_y_reverse() +
        theme_minimal() +
        labs(title=paste0(pos_id, " - Nucleus (feature only)"))
    }
    
    
    p3 <- feat_df %>% 
      dplyr::mutate(feature_id = if_else(type!="nucleolus", "other_feature", feature_id)) %>%
      # dplyr::pull(feature_id) %>% table()
      plot_outline_topView(color_by="feature_id", 
                           color_map=nucleo_col_map,
                           line_alpha=line_alpha, line_width=line_width) +
      coord_fixed(ratio=1, xlim=c(lim_list$xmin, lim_list$xmax), ylim=c(lim_list$ymin, lim_list$ymax)) +
      scale_y_reverse() +
      labs(title=paste0(pos_id, " - Nucleolus (all)"))
    
    plot_df <- feat_df %>% 
      dplyr::filter(type=="nucleolus", str_detect(feature_id, "^nuc"))
    if(nrow(plot_df) > 0){
      p4 <- plot_df %>% 
      plot_outline_topView(color_by="feature_id", 
                           color_map=nucleo_col_map,
                           line_alpha=line_alpha, line_width=line_width) +
      coord_fixed(ratio=1, xlim=c(lim_list$xmin, lim_list$xmax), ylim=c(lim_list$ymin, lim_list$ymax)) +
      labs(title=paste0(pos_id, " - Nucleolus (feature only)"))
    }else{
      p4 <- ggplot() +
        coord_fixed(ratio=1, xlim=c(lim_list$xmin, lim_list$xmax), ylim=c(lim_list$ymin, lim_list$ymax)) +
        scale_y_reverse() +
        theme_minimal() +
        labs(title=paste0(pos_id, " - Nucleolus (feature only)"))
    }
  })
  
  p_grid <- plot_grid(p1, p2, p3, p4, nrow=2, ncol=2, byrow=FALSE)
  ggsave(plot=p_grid, filename=paste0(plot_dir, "/", pos_id, "_outline.png"), width=10, height=10, units="in", dpi=300)
  
  if(F){
    feat_coord_df %>% 
      dplyr::filter(type == "nucleus") %>% 
      plot_ly(data=.,
              x = ~x, y = ~y, z = ~z,
              # type = 'mesh3d', split = ~roi, facecolor= NA,
              type = "scatter3d", mode="markers", color=~feature_id, colors=feature_col_map,
              size=0.5, alpha=0.5)
  }
  
}
```


```{r, include=FALSE}
all_res_df %>% 
  left_join(., dplyr::select(pos_info, pos=position, everything()), by="pos") %>% 
  dplyr::filter(type == "nucleus", ch==1) %>% 
  mutate(valid_feature = str_detect(feature_id, "^nuc")) %>% 
  ggplot(aes(x=circ, y=area)) +
  facet_grid(celltype~valid_feature) +
  geom_point(alpha=0.1)
```


# Comparing signal

***************************************************************************************************

## Per feature signal

```{r}
# channel_map <- c("DAPI", "H3K27ac", "Nr5a2")
feature_res_df <- all_res_df %>% 
  dplyr::filter(str_detect(feature_id, "^nuc"), !is.na(nucleus_id)) %>% 
  mutate(channel = channel_map[as.integer(ch)]) %>% 
  # signal per ROI
  mutate(signal=intden/area) %>% 
  # dplyr::select(signal, mean)
  group_by(pos, type, nucleus_id, channel) %>% 
  reframe(n_z = length(z),
          # IntDen
          sum_intDen = sum(intden),
          mean_intDen = mean(intden),
          median_intDen = median(intden),
          sd_intDen = sd(intden),
          # Area
          sum_area = sum(area),
          mean_area = mean(area),
          median_area = median(area),
          sd_area = sd(area),
          # Signal
          mean_signal = mean(signal),
          median_signal = median(signal),
          sd_signal = sd(signal, na.rm=TRUE)) %>% 
  # Re-calculate the signal
  mutate(signal = sum_intDen/sum_area) %>% 
  left_join(., dplyr::select(pos_info, pos=position, everything()), by="pos")

feature_res_df %>%
  ggplot(aes(x=mean_signal, y=signal)) +
  geom_abline() +
  geom_point(alpha=0.5) +
  coord_fixed()
```

### Raw signal per nucleus
```{r, fig.width=7, fig.asp=0.7}
feature_res_df %>% 
  dplyr::filter(type=="nucleus") %>% 
  ggplot(aes(x = condition, y=median_signal)) +
  geom_boxplot(outlier.alpha=0) +
  geom_quasirandom(aes(size = sd_signal), alpha=0.2) +
  facet_grid(channel~celltype, scales="free_y") +
  labs(title = "Nucleus") +
  theme_bw()

feature_res_df %>% 
  dplyr::filter(type=="nucleolus") %>% 
  ggplot(aes(x = condition, y=median_signal)) +
  geom_boxplot(outlier.alpha=0) +
  geom_quasirandom(aes(size = sd_signal), alpha=0.2) +
  facet_grid(channel~celltype, scales="free_y") +
  labs(title = "Nucleolus") +
  theme_bw()
```

## Per embryo signal
```{r}
embryo_res_df <- feature_res_df %>% 
  group_by(pos, type, channel) %>% 
  reframe(n = length(nucleus_id),
          median_area = median(mean_area),
          sd_area = sd(mean_area),
          mean_area = mean(mean_area),
          mean_signal = mean(signal),
          median_signal = median(signal),
          sd_signal = sd(signal)) %>% 
  left_join(., dplyr::select(pos_info, pos=position, everything()), by="pos")
```

```{r}
embryo_res_df %>% 
  dplyr::select(pos, type, sd_area, mean_area) %>% 
  unique() %>% 
  dplyr::filter(type == "nucleus") %>% 
  left_join(., dplyr::select(pos_info, pos=position, everything()), by="pos") %>% 
  ggplot(aes(x=condition, y=mean_area)) +
  geom_boxplot() +
  geom_quasirandom(aes(size=sd_area), alpha=0.5) +
  facet_grid(.~celltype) +
  labs(title="Cross section area of nucleus") +
  theme_bw()
```


```{r, fig.width=7, fig.asp=0.7}
embryo_res_df %>% 
  dplyr::filter(type == "nucleus") %>% 
  ggplot(aes(x = condition, y=mean_signal)) +
  geom_boxplot(outlier.alpha=0) +
  geom_quasirandom(alpha=0.5) +
  facet_grid(channel~celltype, scales="free_y") +
  labs(title = "Nucleus (Raw signal)") +
  theme_bw()

embryo_res_df %>% 
  dplyr::filter(type == "nucleolus") %>% 
  ggplot(aes(x = condition, y=mean_signal)) +
  geom_boxplot(outlier.alpha=0) +
  geom_quasirandom(alpha=0.5) +
  facet_grid(channel~celltype, scales="free_y") +
  labs(title = "Nucleolus (Raw signal)") +
  theme_bw()
```


# Normalized with background data

```{r}
feature_res_df_norm <- feature_res_df %>% 
  dplyr::select(pos, type, nucleus_id, channel, n_z, sum_intDen, mean_intDen, sum_area, mean_area, feature_signal=signal) %>% 
  left_join(background_df_2, by=c("pos", "channel")) %>% 
  mutate(norm_bg_signal = mean_intDen - (mean_area*bg_mean_signal), 
         norm_cyto_signal = mean_intDen - (mean_area*cyto_mean_signal)) %>% 
  left_join(., dplyr::select(pos_info, pos=position, everything()), by="pos")
```

```{r}
feature_res_df_norm %>% 
  dplyr::filter(type == "nucleus", include) %>% 
  ggplot(aes(x = condition, y=norm_bg_signal)) +
  geom_boxplot() +
  geom_quasirandom(alpha=0.5) +
  facet_grid(channel~celltype, scales="free_y") +
  labs(title = "Per nucleus norm. signal (background)") +
  theme_bw()

feature_res_df_norm %>% 
  dplyr::filter(type == "nucleus", include) %>% 
  ggplot(aes(x = condition, y=norm_cyto_signal)) +
  geom_boxplot() +
  geom_quasirandom(alpha=0.5) +
  facet_grid(channel~celltype, scales="free_y") +
  labs(title = "Per nucleus norm. signal (cytoplasm)") +
  theme_bw()
```


### Per embryo signal
```{r}
embryo_res_df_norm <- feature_res_df_norm %>% 
  dplyr::filter(type == "nucleus", include) %>% 
  group_by(pos, type, channel) %>% 
  reframe(n=length(pos),
          mean_norm_bg_signal = mean(norm_bg_signal),
          sd_norm_bg_signal = sd(norm_bg_signal),
          mean_norm_cyto_signal = mean(norm_cyto_signal),
          sd_norm_cyto_signal = sd(norm_cyto_signal)) %>% 
  left_join(., dplyr::select(pos_info, pos=position, everything()), by="pos") %>% 
  # reorder stuffs by making it a factor
  mutate(channel_factor = factor(channel, levels=channel_order),
         condition_factor = factor(condition, levels=condition_order))
```

```{r}
embryo_res_df_norm %>% 
  dplyr::filter(channel == "DAPI") %>% 
  group_by(celltype, condition) %>% 
  reframe(n=length(pos))
```


```{r, fig.width=8, fig.asp=1}
compare_list <- list(c("fl/fl", "KO"))


p1 <- embryo_res_df_norm %>% 
  ggplot(aes(x = condition_factor, y=mean_norm_bg_signal)) +
  geom_hline(yintercept=0, color="gray25", linetype=2) +
  geom_boxplot(outlier.alpha=0) +
  geom_quasirandom(aes(color=genotype), alpha=0.5) +
  scale_color_manual(values=genotype_color_map) +
  facet_grid(channel_factor~celltype, scales="free_y") +
  # facet_grid(channel_factor~celltype) +
  stat_compare_means(comparisons=compare_list, method="wilcox.test", vjust=1) +
  labs(title = "Per embryo signal (Background norm.)", 
       subtitle = "mean_intDen - (mean_area * mean_background_signal)", 
       x= "", y="Fluorescent Signal (A.U.)") +
  theme_bw() +
  theme(panel.grid=element_blank(), 
        strip.text.y=element_text(size=14, angle=0),
        strip.background.y=element_blank(),
        legend.position="none")


p2 <- embryo_res_df_norm %>% 
  ggplot(aes(x = condition_factor, y=mean_norm_cyto_signal)) +
  geom_hline(yintercept=0, color="gray25", linetype=2) +
  geom_boxplot(outlier.alpha=0) +
  geom_quasirandom(aes(color=genotype), alpha=0.5) +
  scale_color_manual(values=genotype_color_map) +
  facet_grid(channel_factor~celltype, scales="free_y") +
  labs(title = "Per embryo signal (Cytoplasm norm.)", 
       subtitle = "mean_intDen - (mean_area * mean_background_signal)", 
       x= "", y="Fluorescent Signal (A.U.)") +
  stat_compare_means(comparisons=compare_list, method="wilcox.test", vjust=1) +
  theme_bw() +
  theme(panel.grid=element_blank(), 
        strip.text.y=element_text(size=14, angle=0),
        strip.background.y=element_blank(),
        legend.position="none")

print(p1)
print(p2)
```

```{r, fig.width=10, fig.asp=1}
# ggplotly(p2)
```

# Check feature
```{r}
all_pos_df <- tibble(pos = paste0("Position", str_pad(pos_range, width=3, side="left", pad="0")))
n_feature_count_df <- embryo_res_df %>% 
  dplyr::select(pos, type, n) %>% 
  unique() %>% 
  left_join(all_pos_df, ., by="pos") %>% 
  mutate(type = if_else(!is.na(type), type, "nucleus")) %>% 
  tidyr::complete(pos, type, fill=list(n=0)) %>% 
  mutate(pos_id = as.numeric(str_remove_all(str_extract(pos, "\\d+$"), "^0+"))) %>% 
  # Add position info
  left_join(., dplyr::select(pos_info, pos=position, everything()), by="pos")

```

Reporting problematic position

```{r}
n_feature_count_df %>% 
  dplyr::filter(type=="nucleus", n==0) %>% 
  dplyr::select(pos, condition, n)
```

```{r}
n_feature_count_df %>% 
  dplyr::filter(type == "nucleus") %>% 
  ggplot(aes(x=pos_id, y=n, color=celltype)) +
  geom_hline(yintercept=0, color="gray25") +
  # geom_col() +
  geom_point() +
  scale_y_continuous(breaks=seq(0, max(n_feature_count_df$n))) +
  theme_minimal()
```

